---
title: "Analysis for Pascoal et al. 2022"
author: "R. Mashoodh"
date: "02/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(survival)
library(survminer)
#library(patchwork)
#library(kableExtra)
library(gtsummary)

```

### Testing predictions 1 & 2

_**(1) that the harsh conditions of the No Care environment eliminates any genetic variants that are not adapted to survive under these conditions.  
(2) That the more benign conditions of the Full Care environment relax selection by promoting the survival of all genetic variants.**_  

We can do this by focusing on generation 1 and a Binomial GLM to predict the probability of brood success.

Here we define success as brood success (i.e., Brood mass > 0g). Note that we switch to lineage success in the survival analysis.

#### First we plot raw survival proportions:

I've split by Block as this becomes slightly significant later.

```{r, echo=F}

load("data0.RData")

data0<- data0 %>% 
  mutate(LineageID = paste(Male,Female,Breeding,Pop,sep="_")) %>% 
  rename(Evolved = Cond1,
         CurrentEnv = Cond2) %>% 
  mutate(Cond = paste(Evolved, CurrentEnv, sep=""))

gen2 <- data0 %>% 
  filter(Gen == "2")

gen1 <- data0 %>% 
  filter(Gen == 1) %>% 
  mutate(survival = ifelse(Total.weight > 0, 1, 0), #Brood success
         survival2 = ifelse(LineageID %in% gen2$LineageID, 1, 0)) #Defining survival as lineage

```


```{r}
binom.full <- glm(survival ~ Evolved*CurrentEnv + Breeding + as.numeric(Block) + Carcass.weight, data=gen1, family="binomial")

tbl_regression(binom.full) %>%
  as_flex_table() %>% 
  flextable::save_as_docx(., path = "full_binom.docx")

```


#### Binomial GLM for Inbreeding:

```{r, include=F}

## Gen1 -- Inbreeding
gen1.binom.in <- glm(survival ~ Evolved + CurrentEnv + as.numeric(Block) + Carcass.weight, data=subset(gen1, Breeding=="in"), family="binomial")

summary(gen1.binom.in)

```

Pairwise comparisons

```{r}

## Pop diffs within a NC env
gen1.binom.in.NC <- glm(survival ~ Evolved + Block + Carcass.weight, 
                     data=subset(gen1, Breeding=="in" & CurrentEnv == "N"), 
                     family="binomial")


summary(gen1.binom.in.NC)

tbl_regression(gen1.binom.in.NC)

## Pop diffs within a FC env
gen1.binom.in.FC <- glm(survival ~ Evolved + Block + Carcass.weight, 
                     data=subset(gen1, Breeding=="in" & CurrentEnv == "F"), 
                     family="binomial")


summary(gen1.binom.in.FC)

tbl_regression(gen1.binom.in.FC)
```



#### Binomial GLM for Outbreeding


```{r, include = F}
gen1.binom.out <- glm(survival ~ Evolved + CurrentEnv + as.numeric(Block) + Carcass.weight, data=subset(gen1, Breeding=="out"), family="binomial")

summary(gen1.binom.out)


# library(sjPlot)
# tab_model(gen1.binom.out)

```

# Make a table with 2 regressions

```{r}

tbl.in <- tbl_regression(gen1.binom.in)
tbl.out <- tbl_regression(gen1.binom.out)

tbl_merge(
  tbls = list(tbl.in, tbl.out),
  tab_spanner = c("**Inbred**", "**Outbred**")) %>%
  as_flex_table() %>% 
  flextable::save_as_docx(., path = "binomial_tables.docx")


```




#### Figure 1

```{r, echo = F}
df.pred <- data.frame(Evolved = c("F", "F", "N", "N"),
           CurrentEnv = c("F", "N", "F", "N"),
           Carcass.weight = mean(gen1$Carcass.weight),
           Block = c(1.5))

in.pred <- cbind(df.pred,
      data.frame(predict(gen1.binom.in, df.pred, type="response", se.fit =T)))

out.pred <- cbind(df.pred,
      data.frame(predict(gen1.binom.out, df.pred, type="response", se.fit =T)))


all.pred <- rbind(in.pred,out.pred)
all.pred$Breeding <- c(rep("Inbred",4), rep("Outbred",4))


labs <- c(expression(FC['POP']),
            expression(NC['POP']))

all.pred %>% 
  mutate(CurrentEnv = paste0(CurrentEnv,"C")) %>% 
  ggplot(., aes(x=CurrentEnv, y=fit, color=Evolved, group=Evolved)) +
  geom_point() +
  geom_pointrange(aes(ymin = fit - se.fit, ymax = fit + se.fit)) +
  geom_line() +
  ylim(0,1) +
  facet_wrap(.~Breeding) +
  ylab("Survival Probability (Brood Success)") +
  xlab("Current Environment") +
  scale_color_manual(values=c("#e41a1c", "#377eb8"), labels = labs) +
  guides(color=guide_legend("Population")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank())

```


## Figure 2 

Part A - Inbred survival

```{r}
in.dat <- data0 %>% 
  filter(Breeding == "in") %>% 
  group_by(LineageID) %>% 
  filter(Gen == max(Gen))

in.dat$sr <- 1

fits <- survfit(Surv(Gen, sr)~Cond, data=in.dat)

print(fits)

# med <- survminer::surv_median(fits) 
# 
# med %>% 
#   mutate(group = "c") %>% 
#   ggplot(., aes(x=strata, y=median, color=strata)) +
#   geom_point() 

#summary(fits)

cols = c("#e41a1c", "#e41a1c", "#377eb8", "#377eb8")
lines = c("solid", "dotted", "dotted", "solid")

ggsurvplot(
  fit = fits,
  xlab = "Generation",
  ylab = "Survival Probability",
  break.time.by = 1,
  conf.int = T, 
  palette=cols,
  linetype=lines, 
  size=0.75,
  legend="bottom",
  legend.title = "none")

```

Part B - Inbreeding Depression

```{r}
data0$Cond <- with(data0, paste0(Evolved, CurrentEnv))

Conditions <- c("FF", "FN", "NF", "NN")

dep.r <- NULL

for(i in 1:length(Conditions)){
  
  test <- data.frame(with(subset(data0, Cond==Conditions[i]), tapply(Larvae, list(Gen, Breeding), sum)))
  
  dep.r <- cbind(dep.r, (test[,2]-test[,1])/test[,2])
  
}

dep.r <- data.frame(dep.r)
dep.r$Generation <- c(1:8) 
names(dep.r) <- Conditions


dep.r %>% 
  pivot_longer(1:4, names_to = "Cond") %>% 
  mutate(Evolved = substr(Cond,1,1), 
         CurrentEnv = substr(Cond,2,2),
         Native = ifelse(Evolved == CurrentEnv, 1, 0)) %>% 
  ggplot(., aes(x=Generation, y=value, color=Evolved, linetype=as.factor(Native))) +
  geom_point(size = 2) +
  geom_line(size = 0.8) +
  scale_x_continuous(breaks = seq(1,8,by = 1)) +
  scale_color_manual(values=c("#e41a1c", "#377eb8")) +
  scale_linetype_manual(values=c("dotted", "solid")) +
  theme_classic() +
  ylim(-0.25,1) +
  ylab(expression(paste("Inbreeding depression ", "(", delta, ")"))) +
  theme(legend.position = "none")

```


Cox Models & Forest plots


```{r}

in.surv.mod <- coxph(Surv(Gen, sr)  ~ Evolved + CurrentEnv + Block, 
                data=in.dat)

ggforest(in.surv.mod, data=in.dat)

```


#### Outbreeding Survival Analysis

```{r, echo=F}

Time <- population <- NULL

data00 <- data0

for(k in 1:8){

pop <- unique(data00$Pop)[k]

G1Pair <- subset(data00, Pop==pop&Gen==1&Breeding=="out")$Pair



family.list <- list()

for(j in 1:length(G1Pair)){

pair <- list()

pair[[1]] <- G1Pair[j]

for(i in 2:8){

pair[[i]] <- subset(data00, Gen==i&(Male%in%pair[[i-1]]|Female%in%pair[[i-1]])&Pop==pop&Breeding=="out")$Pair

}

family.list[[j]] <- pair

}



times <- unlist(lapply(lapply(family.list, function(x)lapply(x, length)), function(x)sum(x!=0)))

Time <- c(Time, times)

population <- c(population, rep(pop, length(times)))

}



data.s.out <- data.frame(Time=Time, Pop=population)

data.s.out$sr <- 1

data.s.out[data.s.out$Time==8, "sr"] <- 0

data.s.out$sTime <- with(data.s.out, Surv(Time, sr))



data.s.out$Evolved <- substr(data.s.out$Pop, 1, 1)

data.s.out$CurrentEnv <- substr(data.s.out$Pop, 3, 3)

data.s.out$Block <- substr(data.s.out$Pop, 2, 2)

data.s.out$Cond <- as.factor(with(data.s.out, paste0(Evolved, CurrentEnv)))

out.surv.mod <- coxph(sTime  ~ Evolved + CurrentEnv + Block, 
                data=data.s.out)

#summary(out.surv.mod)

#summary(in.surv.mod)

ggforest(out.surv.mod, data=data.s.out)

```

#### Survival plot for Outbreeding:

```{r, echo=F}
fits2 <- survfit(sTime  ~ Cond, data=data.s.out)

#summary(fits2)

ggsurvplot(
  fit = fits2,
  xlab = "Generation",
  ylab = "Survival Probability",
  break.time.by = 1,
  conf.int = T, 
  palette=cols,
  linetype=lines, 
  size=0.75,
  legend="bottom",
  legend.title = "none")

pairwise_survdiff(sTime ~ Cond, data=data.s.out, p.adjust.method = "BH")
```

### Median survival times: Supplementary Table X 

```{r}
in.medians <- surv_median(fits) %>% 
  mutate(Breeding = "Inbred",
         Evolved = c("FC", "FC", "NC", "NC"),
         CurrentEnv = c("FC", "NC", "FC", "NC")) %>% 
  select(Breeding, Evolved, CurrentEnv, median, lower, upper)

out.medians <- surv_median(fits2) %>% 
  mutate(Breeding = "Outbred",
         Evolved = c("FC", "FC", "NC", "NC"),
         CurrentEnv = c("FC", "NC", "FC", "NC")) %>% 
  select(Breeding, Evolved, CurrentEnv, median, lower, upper)


rbind(in.medians, out.medians) %>% 
  flextable::flextable() %>% 
  flextable::save_as_docx(., path = "median_tables.docx")

## median differences in inbred lines

kruskal.test(Gen ~ as.factor(Evolved), data = subset(in.dat, CurrentEnv=="N"))

kruskal.test(Gen ~ as.factor(Evolved), data = subset(in.dat, CurrentEnv=="F"))

```




